Synthesizing Symmetric Lenses

### 2018-07-11 papers/SymmetricLenses/main.tex
add ref to fig 4

However, there is a price for narrowing the synthesis domain so
dramatically: Many real data formats are almost, but not quite,
bijections. More often two related data formats have some data in
common, but also contain some differences. For instance, two related
tables might share four of five columns, but each has a different
fifth column. Alternately, one ad hoc configuration file might include
some file-specific meta data, such as a date, a time or a reference
number, whereas the other does not. Indeed of the 45 benchmark problems
described in Section \ref{sec:evaluation}, all of the benchmarks
taken from Flash Fill~\cite{flashfill} and all of the benchmarks that
synchronize between two ad hoc file formats have this 
characteristic, \afm{num} benchmarks in total. Consequently, tools that search
only for strict bijections will rarely be applicable in practice.

@@

We implemented the \GreedySynth and \Expand algorithms and evaluated their
effectiveness at synthesizing lenses on a 
benchmark suite of 45 problems, including data cleaning tasks taken
from Flash Fill~\cite{gulwani-popl-2014}, information 
extraction tasks \ksf{Dave wants to change information extraction to
soemthing else.} taken from Augeas~\cite{Augeas}, and ad hoc format synchronization tasks
taken from a variety of other locations\footnote{The majority of these format
  synchronization tasks were found from the following index:
  \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.

@@

We have a means of turning stochastic regular expressions into stochastic DNF
regular expressions. The conversion algorithm itself, written
$\ToDNFRegexOf{\Regex}$, is defined below.
\[
  \begin{array}{rcl}
    \ToDNFRegexOf{\String} & = & \DNFOf{(\SequenceOf{\String},1)}\\
    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
  \end{array}
\]
We have proved that this conversion respects languages and probability
distributions.

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
  ad hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used \afm{8} benchmarks from synchronization tasks between a
  variety of real-life formats. These tasks range from synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}

@@

Two classical symmetric lenses are equivalent if they output the same
formats given any sequence of edits. Formally, given a lens
$\Lens \in X \leftrightarrow Y$,
an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
the function $apply$, which, given a lens and an element of that lens's
complement, is a function from sequences of edits to sequences of edits. If
$apply(\Lens,c,es) = es'$, then that means, starting with complement $c$, after
edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
\begin{centermath}
\begin{tabular}[b]{l@{\qquad \; \qquad}l}
$
  \inferrule
  {
  }
  {
    apply(\Lens,c,[]) = []
  }
$
&
$
  \inferrule
  {
    \Lens.putr(x,c) = (y,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
  }
$
\end{tabular}
\end{centermath}
\[
  \inferrule
  {
    \Lens.putl(y,c) = (x,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
  }
\]
\noindent
Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
sequences $es$.

@@

\begin{centermath}
\begin{tabular}[b]{l@{\qquad \; \qquad}l}
$
  \inferrule
  {
  }
  {
    apply(\Lens,xyo,[]) = []
  }
$
&
$
  \inferrule
  {
    \Lens.\CreateROf{x} = y\\
    apply(\Lens,\SomeOf{(x,y)},es) = es'
  }
  {
    apply(l,\None,\InLOf{x}::es) = \InROf{y}::es'
  }
$
\end{tabular}
\end{centermath}
\[
  \inferrule
  {
    \Lens.\CreateLOf{y} = x\\
    apply(\Lens,\SomeOf{(x,y)},es) = es'
  }
  {
    apply(l,\None,\InROf{y}::es) = \InLOf{x}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\PutROf{x'}{y}  = y'\\
    apply(\Lens,\SomeOf{(x',y')},es) = es'
  }
  {
    apply(l,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\PutLOf{y'}{x}  = x'\\
    apply(\Lens,\SomeOf{(x',y')},es) = es'
  }
  {
    apply(l,\SomeOf{(x,y)},\InROf{y'}::es) = \InLOf{x'}::es'
  }
\]

>>>>

However, there is a price for narrowing the synthesis domain so
dramatically: Many real data formats are almost, but not quite,
bijections. More often two related data formats have some data in
common, but also contain some differences. For instance, two related
tables might share four of five columns, but each has a different
fifth column. Alternately, one ad hoc configuration file might include
some file-specific meta data, such as a date, a time or a reference
number, whereas the other does not. Indeed, all of our benchmarks
taken from Flash Fill~\cite{flashfill} and all of our benchmarks that
synchronize between two ad hoc file formats described in
Section \ref{sec:evaluation} \ksf{is this right?} have this 
characteristic. Consequently, past tools that search for strict
bijections will rarely be applicable in practice.

@@

We implemented the \GreedySynth and \Expand algorithms and evaluated their
effectiveness at synthesizing lenses on a 
benchmark suite of 45 problems, including data cleaning tasks taken
from Flash Fill~\cite{gulwani-popl-2014}, information 
extraction tasks \ksf{Dave wants to change information extraction to
soemthing else.} taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
taken from a variety of other locations\footnote{The majority of these format
  synchronization tasks were found from the following index:
  \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.

@@

We have a means of turning stochastic regular expressions into stochastic DNF
regular expressions. The conversion algorithm itself, written
$\ToDNFRegexOf{\Regex}$, is defined below.
\[
  \begin{array}{rcl}
    \ToDNFRegexOf{\String} & = & \DNFOf{(\SequenceOf{\String},1)}\\
    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
  \end{array}
\]
This conversion utilizes operators defined in
Figure~\ref{fig:dnf-regex-functions}. We have proved that this conversion
respects languages and probability distributions.

@@

\subsection{Benchmark Suite}
Our benchmarks are drawn from three different sources.
\begin{enumerate}
\item We adapted \afm{8} data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
  ad-hoc data present in Linux configuration files into a structured, dictionary
  representation.
\item We used \afm{8} benchmarks from synchronization tasks between a
  variety of real-life formats. These tasks range from synchronizing REST
  and JSON web resource descriptions to synchronizing BibTeX and EndNote
  citation descriptions.
\end{enumerate}

@@

Two classical symmetric lenses are equivalent if they output the same
formats given any sequence of edits. Formally, given a lens
$\Lens \in X \leftrightarrow Y$,
an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
the function $apply$, which, given a lens and an element of that lens's
complement, is a function from sequences of edits to sequences of edits. If
$apply(\Lens,c,es) = es'$, then that means, starting with complement $c$, after
edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
\[
  \inferrule
  {
  }
  {
    apply(\Lens,c,[]) = []
  }
\]
\[
  \inferrule
  {
    \Lens.putr(x,c) = (y,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
  }
\]
\[
  \inferrule
  {
    \Lens.putl(y,c) = (x,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
  }
\]
\noindent
Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
sequences $es$.

@@

\[
  \inferrule
  {
  }
  {
    apply(\Lens,xyo,[]) = []
  }
\]
\[
  \inferrule
  {
    \Lens.\CreateROf{x} = y\\
    apply(\Lens,\SomeOf{(x,y)},es) = es'
  }
  {
    apply(l,\None,\InLOf{x}::es) = \InROf{y}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\CreateLOf{y} = x\\
    apply(\Lens,\SomeOf{(x,y)},es) = es'
  }
  {
    apply(l,\None,\InROf{y}::es) = \InLOf{x}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\PutROf{x'}{y}  = y'\\
    apply(\Lens,\SomeOf{(x',y')},es) = es'
  }
  {
    apply(l,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\PutLOf{y'}{x}  = x'\\
    apply(\Lens,\SomeOf{(x',y')},es) = es'
  }
  {
    apply(l,\SomeOf{(x,y)},\InROf{y'}::es) = \InLOf{x'}::es'
  }
\]
